Random Matrix Theory Tools for the Predictive Analysis of Functional Magnetic Resonance Imaging Examinations

### 2023-04-03 paper/paper/SPIE/response_letter.tex

\documentclass[10pt,letter]{article}
\usepackage[utf8]{inputenc}
\usepackage{csquotes}
\usepackage[british]{babel}
\usepackage{url}
\usepackage{graphicx}
\graphicspath{ {figures/} }
\usepackage{geometry}
\geometry{
  lmargin=2cm,
  rmargin=2cm,
  bmargin=2cm,
}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[
    backend=biber,
    style=authoryear,
    natbib=true,
    sortlocale=en_US,
    url=false,
    doi=true,
    eprint=false
]{biblatex}
\addbibresource{report.bib}
\usepackage{fancyhdr}
\usepackage{color}
\definecolor{darkblue}{rgb}{0.,0.,0.4}
\definecolor{darkred}{rgb}{0.5,0.,0.}
\usepackage[pdftex,colorlinks=true,linkcolor=darkblue,citecolor=darkred,urlcolor=blue]{hyperref}
\usepackage{xcolor}

\newcommand{\longcomment}[1]{}

\begin{document}
% \bibliographystyle{apa}
\fancypagestyle{nofooter}{\fancyfoot{}}
\thispagestyle{nofooter}  %  https://tex.stackexchange.com/questions/136384/remove-footer-without-removing-the-header
\setlength{\headheight}{30.0pt}
\fancyhead[C]{\includegraphics[height=0.5in]{stfx_text.png}}
\fancyhead[L]{\includegraphics[height=0.5in]{crest.png}}
\fancyhead[R]{
    \sffamily
    \scriptsize
    4130 University Avenue \\
    B2G 2W5 \\
    Antigonish, Nova Scotia \\
    Canada
}
\normalsize

\title{Cover Letter}
% \begin{center}
% \includegraphics[width=4in]{stfx_logo.png}
% \end{center}

\longcomment{
Specific comments:

% 1. Lines 42-45: I think it's unlikely that most readers of SPIE will be
% familiar with RMT and so it would be a good idea to have a bit more of an
% introduction - what's a random matrix?, what does RMT try to do (e.g.,
% characterize systems that can be described using matrices, statistically,
% rather than deterministically.) Later they talk about what RMT has been applied
% to, but it would be good to orient the reader to the general idea.

% 2. Line 44: "independently" should be "independent"

% 3. Line 56: This is a pretty broad statement, because it would seem that it's
% an open question whether all complex systems would lend themselves to RMT-based
% analysis.

% 4. Line 100: Do the authors really mean "eliminate" here? Do they instead mean
% "mask", given that the information presumably doesn't disappear.

% 5. Line 121: "functional connectivity". Do the authors mean "functional
% connectivity matrix"?

% 6. Line 138: This sentence is a bit garbled. Do the authors mean "often poorly
% documented or even missing..."?

% 7. Lines 143-145: I don't think these absences raise doubts about these two
% things in general, but rather that one has to carefully document the workflows
% and make the data and code available to the public.

% 8. Line 156: Do the authors mean "allow a reasonable classification of each
% scan without resorting to other information". Unclear what they mean by
% "entirety".

% 9. Table 1: Volumes and n_scans are not defined sufficiently. Are "Volumes" the
% number of volumes per run? Is n_scans to total number of runs? Is there one run
% per participant? The participant info is in Table 2, but Table 1 should be
% independently clear.

% 10. Line 180: "learning vs. rest" should be in a different font to be
% consistent with the other dataset descriptions.

% 11. Line 208: Are the "fractional anisotropy" values mentioned here the input
% variables for BET which govern how aggressive BET is and the gradient
% parameter?

12. Section 3.2: While I understand that the authors are trying to make a
general statement that will apply to all features, this section is unclear
because the reader has no idea at this point what features are. Perhaps this
could be expanded.

% 13. Line 223: How about "computing time" and "storage space", assuming that is
% what is meant.

% 14. Line 267: It's not clear that this section should be here, maybe at the end
% of the methods section?

% 15. Line 292: This is a minor point, but sentences shouldn't start with an
% abbreviation (E.g.) better "For example,". This also occurs in line 387.

% 16. Line 311: I see what they mean by "voxel time series" because of lines
% 317-318, so it might be clearer to move those lines to directly after line 311,
% so it is clear without having to read further.

17. The authors do a nice job of presenting the relevant results, but the
conclusions were rather general, being along the lines of "investigate the
sensitivity of choices made in the processing and analysis". That's an
important result, given previous work and shows that RMT might not "work" with
different data sets and processing workflows. The question is: what's the
takeaway: that the method is not one in which you can have a default set of
parameters that will generally give good results (like say, the recon-all
pipeline for FreeSurfer)? If RMT so sensitive to preprocessing and choice of
parameters, is this a useful method to be applied widely on fMRI data? It would
be good for this to be discussed and perhaps some recommendations given.
}

\noindent
\today, \\
Dear Dr. Gimi, \\

Thank you for your careful consideration and comments on our manuscript
\emph{Random Matrix Theory Tools for the Predictive Analysis of Functional
Magnetic Resonance Imaging Examinations} submitted for publication in SPIE
Journal of Medical Imaging.

In response to reviewer comments, we have made a number of changes to the
manuscript. Minor changes relating to wording or for which an edit is
straightforward (changing or adding a word or two) are identified in
\textcolor{orange}{orange} in the updated \.pdf manuscript (or with
\verb|\textcolor{orange}{text}| in the \.tex file). Other more significant
changes requiring the addition (or movement) of 1-3 sentences are identified
in \textcolor{cyan}{cyan} (\verb|\textcolor{cyan}{text}|).

\begin{description}
\item[Comment 1]: An introduction to the basics of Random Matrix Theory (RMT) has been
added to the introduction, to improve the basic accessibility of RMT to readers
\end{description}

Reviewer comments 2, 4, 5, 6, 10, 11, 13, 15, 16 noted easily-correctable
issues relating to phrasing and clarity, and we have made changes in line with
those comments. Reviewer comments 3 and 7 noted some overly-broad or
insufficiently specific claims on our part. We agree these need qualification,
and have altered these statements to be avoid implying anything overly-broad.
Reviewer comments 8, 9 noted some issues with vagueness involving descriptions
of the fMRI scans in the text and tables. Both the text and table captions have
been improved to make these details clear. Comment 14 also suggested moving some
material to methods, and this was also done.

Comment 1 suggested expanding the introduction to Random Matrix Theory (RMT)
somewhat to better cover the basic aspects of RMT, and help with orienting
readers unfamiliar with the area. We have done this expansion, and hope that it
covers the basics both in a way that presumes neither too little nor too much
from readers (always a very difficult balance to strike with something like
RMT, unfortunately).

Our reviewer's final comment suggested perhaps strengthening our conclusions
re: the utility of RMT, given sensitivity to preprocessing.

Please address all correspondence concerning this manuscript to me at \href{jlevman@stfx.ca}{jlevman@stfx.ca}.

\bigskip

\noindent
Sincerely, \\

\small
\noindent
\textbf{Corresponding Author}\\
Jacob Levman, PhD (jlevman@stfx.ca) \\
\footnotesize
\hspace*{0.25cm}St. Francis Xavier University, Department of Computer Science  \\
\hspace*{0.25cm}Martinos Center for Biomedical Imaging, Massachusetts General Hospital \\
\hspace*{0.25cm}Harvard Medical School, Department of Radiology (jlevman@mgh.harvard.edu) \\
\hspace*{0.25cm}Nova Scotia Health Authority, Research Affiliate (jacob.levman@nshealth.ca) \\
\normalsize

\small
\noindent
\textbf{Additional Contacts} [should the corresponding author be unavailable]\\
Derek Berger (dberger@stfx.ca)\\
\footnotesize
\hspace*{0.25cm} St. Francis Xavier University, Department of Computer Science \\
\small
Gurpreet S. Matharoo (gmatharo@stfx.ca) \\
\footnotesize
\hspace*{0.25cm} St. Francis Xavier University, Department of Physics \\
\hspace*{0.25cm} St. Francis Xavier University, ACENET
\normalsize

% \affil[b]{St. Francis Xavier University, ACENET, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}

% \affil[c]{St. Francis Xavier University, Department of Physics, 4130 University Avenue, Antigonish, Nova Scotia, Canada, B2G 2W5}

% \affil[d]{Martinos Center for Biomedical Imaging, 149 Thirteenth Street, Suite 2301, Charlestown, Massachusetts, United States, 02129}
% \affil[e]{Harvard Medical School, Department of Radiology, 25 Shattuck Street Boston, Massachusetts 02115, United States, 02129}
% \affil[f]{Nova Scotia Health Authority, Research Affiliate, Nova Scotia, Canada}

% \printbibliography % biblatex
% \renewcommand{\bibsection}{}  % https://tex.stackexchange.com/questions/132646/how-to-remove-the-references-title
% \smallskip
% \bibliographystyle{mystyle}
% \bibliography{cover}  % natbib

\footnotesize
\begin{center}
\noindent\rule{6cm}{0.4pt}
\end{center}

\AtNextBibliography{\footnotesize}
\printbibliography[heading=none]

\end{document}

### 2023-04-03 paper/paper/SPIE/article.tex
add xcolor dependency
[WIP] make paper corrections

\documentclass[12pt]{spieman}  % 12pt font required by SPIE;
%\documentclass[a4paper,12pt]{spieman}  % use this instead for A4 paper
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\graphicspath{ {./figures/} }
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{tocloft}
\usepackage{doi}
\usepackage{lineno}
\linenumbers
\usepackage{csquotes}
\MakeOuterQuote{"}
\usepackage{xr}
\externaldocument[S-]{supplementary}

@@

At its most rudimentary, RMT describes the expected behavior of the
eigenvalues—also often called levels—of a number of classes of random
matrices\cite{andersonIntroductionRandomMatrices2010,
akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. A random matrix is,
simplifying somewhat, a matrix in which the entries are drawn from
independently and identically distributed (\textit{iid}) distributions. Though
first developed to describe the fluctuation of nuclei energy levels in quantum
physics \cite{mehtaRandomMatrices2004, guhrRandommatrixTheoriesQuantum1998a},
RMT has been shown to have extremely broad potential. In small-scale physical
systems, RMT universalities have been observed in quantum chaotic systems,
complex nuclei, atoms, molecules and disordered mesoscopic systems
\cite{guhrRandommatrixTheoriesQuantum1998a, mehtaRandomMatrices2004,
brodyRandommatrixPhysicsSpectrum1981, beenakkerRandommatrixTheoryQuantum1997,
bohigasHigherOrderCorrelationsSpectra1985, wintgenLevelStatisticsQuantized1988,
pandeySkewOrthogonalPolynomialsUniversality2001}, and at larger scales, RMT has
been applied to atmospheric physics
\cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations
\cite{plerouRandomMatrixApproach2002}, social networks
\cite{jalanUncoveringRandomnessSuccess2014}, random networks
\cite{bandyopadhyayUniversalityComplexNetworks2007}, network-formation in
liquids\cite{sastrySpectralStatisticsInstantaneous2001,
matharooSpectralStatisticsQuenched2009} and amorphous clusters
\cite{sarkarUniversalityVibrationalSpectra2004,
matharooVibrationalSpectraAmorphous2005,
matharooUniversalityVibrationalSpectra2008}. Within biological systems, RMT has
also been used to successfully model aspects of  amino acid functional
relationships \cite{bhadolaTargetingFunctionalMotifs2016}, synchrony in
epileptic seizures \cite{osorioPhasesynchronizationRandommatrixBased2011}, and
in protein-protein interactions both in different species
\cite{agrawalQuantifyingRandomnessProtein2014} and breast cancer
\cite{raiRandomnessPreservedPatterns2015}. RMT has also been used to guide
statistical decisions in principal components
analyses\cite{franklinParallelAnalysisMethod1995,
veraartDenoisingDiffusionMRI2016, ulfarssonDimensionEstimationNoisy2008}, and,
more recently, has provided insights into the behaviors of deep neural
networks\cite{martinImplicitSelfRegularizationDeep2021,
martinPredictingTrendsQuality2021}.

@@

RMT has also been used in ROI-based fMRI functional connectivity studies to
investigate differences between rest and task states
\cite{wangSpectralPropertiesTemporal2015}, between subjects with and without
attention-deficit hyperactive disorder (ADHD)
\cite{wangRandomMatrixTheory2016}, between pain and non-pain states
\cite{matharooSpontaneousBackpainAlters2020}, and between left-sided vs.
right-sided motor imagery \cite{guRandomMatrixTheory2020}. While no differences
were found in the latter study, across the first three studies, the spectra of
resting or low-attention states exhibited properties close to the GOE. These
findings suggest that certain aspects of psychological processes might be
characterized, in part, by features computed from the eigenvalues of fMRI
correlation matrices, and that these features might vary in an interpretable
manner across psychological processes. If this is the case, RMT could aid in
understanding the functioning of the human brain.

@@

However, from a predictive standpoint, these summary statistics may eliminate
predictively-useful information. If the basic RMT insight is that the
eigenvalues alone can provide understanding of a system, then those eigenvalues
(or other simple transformations of them) ought also to be predictively useful.
We examine a number of such features in this study, and refer to both
RMT-derived features like the rigidity and level variance, and non-RMT-derived
features collectively as \textit{eigenfeatures} in this study.

@@

RMT suggests a potentially useful reduction in the form of the eigenvalues of
the voxelwise functional connectivity. An \(N \times t\) matrix of
\(N\) time series of length \(t\), and where \(t \ll N\) will have a symmetric
correlation (or covariance) matrix with \(t - 1\) non-zero, positive
real-valued eigenvalues. These eigenvalues can be computed highly
efficiently via transposition of the voxelwise correlation or covariance
matrix.

@@

In addition, the previous papers do not provide publicly-accessible code to
reproduce results. The extraction of the spectral rigidity and level variance
is computationally-demanding and mathematically and algorithmically
non-trivial, and additionally requires an \textit{unfolding}
procedure\cite{guhrRandommatrixTheoriesQuantum1998a, mehtaRandomMatrices2004}.
Unfolding is an exponential fitting procedure that involves a number of highly
subjective decisions regarding outliers and the flexibility of the fitting
function. These decisions are, unfortunately, often poorly or even entirely
undocumented, but are known to often dramatically impact RMT conclusions
\cite{abul-magdUnfoldingSpectrumChaotic2014,
abueleninSpectralUnfoldingChaotic2018, fossionRandommatrixSpectraTime2013,
abueleninEffectUnfoldingSpectral2012, moralesImprovedUnfoldingDetrending2011}.

@@

The \textit{Learning} dataset\cite{schapiroHumanHippocampalReplay2018,
schapiroHumanHippocampalReplay2020} has both rs-fMRI and task fMRI scans
available for all subjects, and \code{task v rest} is chosen as the binary
classification task for our analysis.  The task is complex and difficult to
summarize here adequately, but involved multiple phases where subjects were
presented with various related abstract images, and then later tested for their
memory of certain aspects of the learned images. Thus, the classification task
might also be considered learning vs. rest.

@@

Brain extraction was performed first with BET
\cite{smithFastRobustAutomated2002}, and then with
ANTs\cite{avantsReproducibleEvaluationANTs2011} to clean up residue left behind
by BET. BET results were visually inspected for each image in each dataset, and
BET fractional anisotropy values were modified to ensure brain extraction was
acceptable. Motion-correction was performed with FSL's MCFLIRT
\cite{jenkinsonImprovedOptimizationRobust2002}, and slice time correction
with FSL's slicetimer\cite{jenkinsonFSL2012} tool. Finally, functional
images were registered directly to the MNI152 template via
ANTs\cite{avantsReproducibleEvaluationANTs2011}.

@@

The time and space complexities of computing the full \(N \times N\)
\(\mathbf{M}\) matrix of Pearson correlation coefficients (and subsequent
eigenvalues) are too large to be tractable. However, since transposition does
not change the eigenvalues, we can use the transpose to efficiently compute
these eigenvalues (see Appendix \ref{sec:transpose}). These raw eigenvalues
(feature \code{eigs} in Table \ref{tab:features}), most directly test the idea
that eigenvalues alone are useful functional connectivity features.

@@

The largest, middle, or smallest eigenvalues could be most useful in
characterizing any given system. E.g. if the smallest eigenvalues correspond to
random / noise aspects of a system, but differences in the nature of the system
noise most differentiate between systems, then the smallest eigenvalues may
have the most predictive utility.  Likewise, the \(L\) value in each RMT
eigenfeature defines the degree of locality at which we summarize the spectrum:
with small \(L\)-values, the spectral rigidity summarizes the non-linearity of
eigenvalues that are relatively close to each other in magnitude. At large
values of \(L\), the rigidity summarizes the long-range non-linearity of the
spectrum.

@@

If RMT- or eigenvalue-based features fail to predictively outperform features
that are simple summary statistics of the fMRI data, it is difficult to justify
the greater computational and interpretational complexities of the former. We
compute ten simple statistical reductions of the voxel time series ( "tseries"
features in Table \ref{tab:features}) to help assess the relative value of RMT
features.

The baseline reductions used as statistical summaries were: robust measures of
location (mean, max, min); non-robust measures of location (median,
\(95^{\text{th}}\) percentile, \(5^{\text{th}}\) percentile); non-robust
measures of scale (standard deviation and the range—\code{T-rng} in Table
\ref{tab:features}); and robust measures of scale (interquartile range and
difference between \(95^{\text{th}}\) percentile and \(5^{\text{th}}\)
percentile—\code{T-rrng} in Table \ref{tab:features}). Each statistic reduces
the fMRI data along the voxel dimension, yielding a feature of \(t\) dimensions
(e.g. the ``mean'' feature, \code{T-mean}, is the usual global mean signal).
% Exact details can be found in the
% \href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
% source code}.

@@

It is important, to reduce figure clutter and complexity, to exclude a
classification task or an analytic factor from summaries if there is no
evidence that the classification task is solvable in general, or if, when
restricting the analytic factor to a particular instance, there is no evidence
of solvability. E.g. if a particular trimming procedure were to render all
classification tasks unsolvable, it would be better to note this and exclude
the associated mAUROCs from further visualizations, rather than to have the
mAUROC distributions diluted by the bad procedure.

>>>>

\documentclass[12pt]{spieman}  % 12pt font required by SPIE;
%\documentclass[a4paper,12pt]{spieman}  % use this instead for A4 paper
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{graphicx}
\graphicspath{ {./figures/} }
\usepackage{setspace}
\usepackage{hyperref}
\usepackage{tocloft}
\usepackage{doi}
\usepackage{lineno}
\linenumbers
\usepackage{csquotes}
\MakeOuterQuote{"}
\usepackage{xr}
\externaldocument[S-]{supplementary}
\usepackage{xcolor}

@@

At its most rudimentary, RMT describes the expected behavior of the
eigenvalues—also often called levels—of a number of classes of random
matrices\cite{andersonIntroductionRandomMatrices2010,
akemannOxfordHandbookRandom2011, mehtaRandomMatrices2004}. A random matrix is,
simplifying somewhat, a matrix in which the entries are drawn from
\textcolor{orange}{independent} and identically distributed (\textit{iid})
distributions. Though first developed to describe the fluctuation of nuclei
energy levels in quantum physics \cite{mehtaRandomMatrices2004,
guhrRandommatrixTheoriesQuantum1998a}, RMT has been shown to have extremely
broad potential. In small-scale physical systems, RMT universalities have been
observed in quantum chaotic systems, complex nuclei, atoms, molecules and
disordered mesoscopic systems \cite{guhrRandommatrixTheoriesQuantum1998a,
mehtaRandomMatrices2004, brodyRandommatrixPhysicsSpectrum1981,
beenakkerRandommatrixTheoryQuantum1997,
bohigasHigherOrderCorrelationsSpectra1985, wintgenLevelStatisticsQuantized1988,
pandeySkewOrthogonalPolynomialsUniversality2001}, and at larger scales, RMT has
been applied to atmospheric physics
\cite{santhanamStatisticsAtmosphericCorrelations2001}, stock cross-correlations
\cite{plerouRandomMatrixApproach2002}, social networks
\cite{jalanUncoveringRandomnessSuccess2014}, random networks
\cite{bandyopadhyayUniversalityComplexNetworks2007}, network-formation in
liquids\cite{sastrySpectralStatisticsInstantaneous2001,
matharooSpectralStatisticsQuenched2009} and amorphous clusters
\cite{sarkarUniversalityVibrationalSpectra2004,
matharooVibrationalSpectraAmorphous2005,
matharooUniversalityVibrationalSpectra2008}. Within biological systems, RMT has
also been used to successfully model aspects of  amino acid functional
relationships \cite{bhadolaTargetingFunctionalMotifs2016}, synchrony in
epileptic seizures \cite{osorioPhasesynchronizationRandommatrixBased2011}, and
in protein-protein interactions both in different species
\cite{agrawalQuantifyingRandomnessProtein2014} and breast cancer
\cite{raiRandomnessPreservedPatterns2015}. RMT has also been used to guide
statistical decisions in principal components
analyses\cite{franklinParallelAnalysisMethod1995,
veraartDenoisingDiffusionMRI2016, ulfarssonDimensionEstimationNoisy2008}, and,
more recently, has provided insights into the behaviors of deep neural
networks\cite{martinImplicitSelfRegularizationDeep2021,
martinPredictingTrendsQuality2021}.

@@

RMT has also been used in ROI-based fMRI functional connectivity studies to
investigate differences between rest and task states
\cite{wangSpectralPropertiesTemporal2015}, between subjects with and without
attention-deficit hyperactive disorder (ADHD)
\cite{wangRandomMatrixTheory2016}, between pain and non-pain states
\cite{matharooSpontaneousBackpainAlters2020}, and between left-sided vs. \
right-sided motor imagery \cite{guRandomMatrixTheory2020}. While no differences
were found in the latter study, across the first three studies, the spectra of
resting or low-attention states exhibited properties close to the GOE. These
findings suggest that certain aspects of psychological processes might be
characterized, in part, by features computed from the eigenvalues of fMRI
correlation matrices, and that these features might vary in an interpretable
manner across psychological processes. If this is the case, RMT could aid in
understanding the functioning of the human brain.

@@

However, from a predictive standpoint, these summary statistics may
\textcolor{orange}{mask} predictively-useful information. If the basic RMT
insight is that the eigenvalues alone can provide understanding of a system,
then those eigenvalues (or other simple transformations of them) ought also to
be predictively useful.
We examine a number of such features in this study, and refer to both
RMT-derived features like the rigidity and level variance, and non-RMT-derived
features collectively as \textit{eigenfeatures} in this study.

@@

RMT suggests a potentially useful reduction in the form of the eigenvalues of
the voxelwise functional connectivity \textcolor{orange}{matrix}. An \(N \times
t\) matrix of \(N\) time series of length \(t\), and where \(t \ll N\) will
have a symmetric correlation (or covariance) matrix with \(t - 1\) non-zero,
positive real-valued eigenvalues. These eigenvalues can be computed highly
efficiently via transposition of the voxelwise correlation or covariance
matrix.

@@

In addition, the previous papers do not provide publicly-accessible code to
reproduce results. The extraction of the spectral rigidity and level variance
is computationally-demanding and mathematically and algorithmically
non-trivial, and additionally requires an \textit{unfolding}
procedure\cite{guhrRandommatrixTheoriesQuantum1998a, mehtaRandomMatrices2004}.
Unfolding is an exponential fitting procedure that involves a number of highly
subjective decisions regarding outliers and the flexibility of the fitting
function. These decisions are, unfortunately, often poorly
\textcolor{orange}{documented} or even entirely \textcolor{orange}{missing from
method descriptions, despite being} known to often dramatically impact RMT
conclusions \cite{abul-magdUnfoldingSpectrumChaotic2014,
abueleninSpectralUnfoldingChaotic2018, fossionRandommatrixSpectraTime2013,
abueleninEffectUnfoldingSpectral2012, moralesImprovedUnfoldingDetrending2011}.

@@

The \textit{Learning} dataset\cite{schapiroHumanHippocampalReplay2018,
schapiroHumanHippocampalReplay2020} has both rs-fMRI and task fMRI scans
available for all subjects, and \code{task v rest} is chosen as the binary
classification task for our analysis.  The task is complex and difficult to
summarize here adequately, but involved multiple phases where subjects were
presented with various related abstract images, and then later tested for their
memory of certain aspects of the learned images. Thus, the classification task
might also be considered \textcolor{orange}{\texttt{learning v rest}}.

@@

Brain extraction was performed first with BET
\cite{smithFastRobustAutomated2002}, and then with
ANTs\cite{avantsReproducibleEvaluationANTs2011} to clean up residue left behind
by BET. BET results were visually inspected for each image in each dataset, and
BET fractional \textcolor{orange}{intensity threshold (\texttt{-f}
argument)} values were modified to ensure brain extraction was acceptable.
Motion-correction was performed with FSL's MCFLIRT
\cite{jenkinsonImprovedOptimizationRobust2002}, and slice time correction
with FSL's slicetimer\cite{jenkinsonFSL2012} tool. Finally, functional
images were registered directly to the MNI152 template via
ANTs\cite{avantsReproducibleEvaluationANTs2011}.

@@

The \textcolor{orange}{computing} time and \textcolor{orange}{memory
requirements} for calculating the full \(N \times N\) \(\mathbf{M}\) matrix of
Pearson correlation coefficients (and subsequent eigenvalues) are too large to
be tractable. However, since transposition does not change the eigenvalues, we
can use the transpose to efficiently compute these eigenvalues (see Appendix
\ref{sec:transpose}). These raw eigenvalues (feature \code{eigs} in Table
\ref{tab:features}), most directly test the idea that eigenvalues alone are
useful functional connectivity features.

@@

The largest, middle, or smallest eigenvalues could be most useful in
characterizing any given system. \textcolor{orange}{For example,} if the
smallest eigenvalues correspond to random / noise aspects of a system, but
differences in the nature of the system noise most differentiate between
systems, then the smallest eigenvalues may have the most predictive utility.
Likewise, the \(L\) value in each RMT eigenfeature defines the degree of
locality at which we summarize the spectrum: with small \(L\)-values, the
spectral rigidity summarizes the non-linearity of eigenvalues that are
relatively close to each other in magnitude. At large values of \(L\), the
rigidity summarizes the long-range non-linearity of the spectrum.

@@

If RMT- or eigenvalue-based features fail to predictively outperform features
that are simple summary statistics of the fMRI data, it is difficult to justify
the greater computational and interpretational complexities of the former. We
compute ten simple statistical reductions of the voxel time series ( "tseries"
features in Table \ref{tab:features}) to help assess the relative value of RMT
features.  \textcolor{orange}{That is, each statistic reduces the fMRI data along
the voxel dimension, yielding a feature of \(t\) dimensions (e.g. the ``mean''
feature, \code{T-mean}, is the usual global mean signal).}

The baseline reductions used as statistical summaries were: robust measures of
location (mean, max, min); non-robust measures of location (median,
\(95^{\text{th}}\) percentile, \(5^{\text{th}}\) percentile); non-robust
measures of scale (standard deviation and the range—\code{T-rng} in Table
\ref{tab:features}); and robust measures of scale (interquartile range and
difference between \(95^{\text{th}}\) percentile and \(5^{\text{th}}\)
percentile—\code{T-rrng} in Table \ref{tab:features}).

% Exact details can be found in the
% \href{https://github.com/DM-Berger/random-matrix-fmri/blob/master/code/rmt/enumerables.py#L51-L69}{paper
% source code}.

@@

It is important, to reduce figure clutter and complexity, to exclude a
classification task or an analytic factor from summaries if there is no
evidence that the classification task is solvable in general, or if, when
restricting the analytic factor to a particular instance, there is no evidence
of solvability. \textcolor{orange}{For example,} if a particular trimming
procedure were to render all classification tasks unsolvable, it would be
better to note this and exclude the associated mAUROCs from further
visualizations, rather than to have the mAUROC distributions diluted by the bad
procedure.
